---
title: "Hands On Exercise 1"
---

## Loading of packages

```{r}
pacman::p_load(sf, tidyverse)
```

## Importing of polygon feature data in shapefile format

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPathGazette")
```

## Importing GIS data in kml format

```{r}
preschool = st_read("data/geospatial/preschools-location.kml")
```

## Checking content of A Simple Feature Data Frame

```{r}
st_geometry(mpsz)
```

Note that the print only displays basic information of the feature class such as type of geometry, the geographic extent of the features and the coordinate system of the data

## Working with glimpse()

```{r}
glimpse(mpsz)
```

## Working with head()

```{r}
head(mpsz, n=5)
```

## Plotting the Geospatial Data

```{r}
plot(mpsz)
```

```{r}
plot(st_geometry(mpsz))
```

```{r}
plot(mpsz["PLN_AREA_N"])
```

## Working with projects

Assigning EPSG code to simple feature data frame

```{r}
st_crs(mpsz)
```

Assign correct EPSG code to mpsz dataframe

```{r}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

Checking CSR if correct again

```{r}
st_crs(mpsz3414)
```

## Transforming projection of preschool from wgs84 to svy21

```{r}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

Note: In practice, we need to find out the appropriate project coordinate system to use before performing projection transformation

## Importing and converting aspatial data

```{r}
listings <- read_csv("data/aspatial/listings.csv")
```

```{r}
list(listings) 
```

## Creating simple feature data frame from an aspatial data frame

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(listings_sf)
```

## Geoprocessing with sf package

### Buffering

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```

Calculating area of buffers

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
```

Sum of Base R to derive total land involved

```{r}
sum(buffer_cycling$AREA)
```

## Point-in-polygon count

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

Check summary

```{r}
summary(mpsz3414$`PreSch Count`)
```

list planning subzone

```{r}
top_n(mpsz3414, 1, `PreSch Count`)
```

derive area of each planning subzone

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
```

Compute density

```{r}
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

## EDA

## Distribution of PreSch Density

```{r}
hist(mpsz3414$`PreSch Density`)
```

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
