---
title: "In Class Exercise 8"
author: "Daniel Chng"
date: "6 March 2023"
date-modified: "`r Sys.Date()`"
format: html
execute: 
  eval: true
  echo: true
  message: false
  warning: false
editor: visual
---

# Overview

Geographically weighted regression (GWR) is a spatial statistical technique that takes non-stationary variables into consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable). In this hands-on exercise, you will learn how to build [hedonic pricing](https://www.investopedia.com/terms/h/hedonicpricing.asp) models by using GWR methods. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

# Importing Modules

{r} pacman::p_load(olsrr, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary, readr, dplyr, tidyselect)

# Importing Data

## Geospatial Data

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

## Aspatial Data

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

## Converting Aspatial Data to sf Object

```{r}
condo_resale.sf <- st_as_sf(condo_resale, coords = c("LONGITUDE", "LATITUDE"), crs=4326) %>% st_transform(crs=3414)
```

# Hedonic Pricing Modelling in R

## Simple Linear Regression Method

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)

ggplot(data=condo_resale.sf,aes(x=AREA_SQM, y=SELLING_PRICE)) + geom_point() + geom_smooth(method = lm)
```

## Multiple Linear Regression Method

```{r}
corrplot::corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
                   t1.pos = "td", t1.cex = 0.5, method = "number", type = "upper")
```

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo.mlr)
```

```{r}
condo.mlr1 <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                   PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                   PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK + 
                   PROX_PRIMARY_SCH + PROX_SHOPPING_MALL    + PROX_BUS_STOP + 
                   NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                 data=condo_resale.sf)
ols_regress(condo.mlr1)
```

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```
